### 1. Load and Process Raw Data
```{r}
library(Seurat)
library(SeuratObject)
library(scrubletR)
library(reticulate)
Sys.setenv(RETICULATE_PYTHON = "/Users/ananyashah/opt/anaconda3/envs/scrublet/bin/python")
use_condaenv("scrublet")

folder.names <- c("E10.5_WT", "E12.5_WT1", "E12.5_WT2", "E14.5_WT1", "E14.5_WT2")
data.path <- "/Users/ananyashah/Desktop/Ananya/"
objs <- list()

for (f in folder.names) {
  path <- file.path(data.path, f)
  data <- Read10X(path)
  obj.temp <- CreateSeuratObject(counts = data, project = f)
  obj.temp <- scrublet_R(seurat_obj = obj.temp, python_home = Sys.getenv("RETICULATE_PYTHON"))
  objs[[f]] <- obj.temp
}

obj <- merge(objs[[1]], objs[2:5])
``` 

```{r}
#obj = readRDS('/Volumes/backup-harddrive/2-isolate_neurons_objs.RDS') 
```


### 2. Filter and Normalize
```{r}
#obj <- obj[, sample(colnames(obj), size = 30000, replace = FALSE)]
obj$LowCount <- with(obj@meta.data, nFeature_RNA < 3000 & nCount_RNA < 50000 & nCount_RNA > 9000)
obj$predicted_doublets <- obj$doublet_scores > 0.5
obj$percent.mt <- PercentageFeatureSet(obj, pattern = "^mt-")

obj <- JoinLayers(obj)
cell_mask <- Reduce(intersect, list(
  WhichCells(obj, expression = nFeature_RNA > 700),
  WhichCells(obj, expression = nFeature_RNA < 10000),
  WhichCells(obj, expression = nCount_RNA < 50000),
  WhichCells(obj, expression = percent.mt < 12.5)
))
gene_mask <- rownames(obj)[Matrix::rowSums(obj@assays$RNA@layers$counts > 0) > 8]
obj <- subset(obj, features = gene_mask, cells = cell_mask)
``` 

### 2b. QC Plots (Violin + Scatter) ---
```{r}

VlnPlot(obj, features = "nFeature_RNA", group.by = "orig.ident", pt.size = 0.1) +
  ggtitle("nFeature_RNA per Sample")
VlnPlot(obj, features = "nCount_RNA", group.by = "orig.ident", pt.size = 0.1) +
  ggtitle("nCount_RNA per Sample")


FeatureScatter(obj, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "orig.ident") +
  ggtitle("nCount_RNA vs. nFeature_RNA (by Sample)")


FeatureScatter(obj, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "predicted_doublets") +
  ggtitle("nCount_RNA vs. nFeature_RNA (by Doublet Status)")


VlnPlot(obj, features = "doublet_scores", group.by = "predicted_doublets", pt.size = 0.1) +
  ggtitle("Doublet Scores by Doublet Prediction")


FeatureScatter(obj, feature1 = "percent.mt", feature2 = "nCount_RNA", group.by = "LowCount") +
  ggtitle("percent.mt vs. nCount_RNA (LowCount Flagged)")


VlnPlot(obj, features = "percent.mt", group.by = "unintegrated_clusters", pt.size = 0.1) +
  ggtitle("percent.mt by Unintegrated Cluster")


### 3. PCA and Clustering
```{r}
obj <- NormalizeData(obj)
obj <- FindVariableFeatures(obj)
obj <- ScaleData(obj)
obj <- RunPCA(obj)
obj <- FindNeighbors(obj, dims = 1:30)
obj <- FindClusters(obj, resolution = 2, cluster.name = "unintegrated_clusters")
obj <- RunUMAP(obj, dims = 1:30, reduction.name = "umap.unintegrated")
``` 

### 4. Initial Visualizations
```{r}
DimPlot(obj, reduction = "umap.unintegrated", group.by = "orig.ident")
DimPlot(obj, reduction = "umap.unintegrated", group.by = "unintegrated_clusters", label = TRUE)
FeaturePlot(obj, "percent.mt", reduction = "umap.unintegrated")
VlnPlot(obj, "percent.mt", group.by = "unintegrated_clusters")
DotPlot(obj, features = c("Sox10", "Uchl1", "Snap25", "Wnt1")) + coord_flip()
``` 

### 5. Integration with CCA
```{r}
obj <- IntegrateLayers(
  object = obj, 
  method = CCAIntegration,
  orig.reduction = "pca", 
  new.reduction = "integrated.cca",
  verbose = FALSE
)
obj <- FindNeighbors(obj, reduction = "integrated.cca", dims = 1:30)
obj <- FindClusters(obj, resolution = 2, cluster.name = "cca_clusters")
obj <- RunUMAP(obj, reduction = "integrated.cca", dims = 1:30, reduction.name = "umap.cca")
``` 

```{r}
obj = readRDS("/Volumes/backup-harddrive/Ananya/development_all_integrated.rds")
```

### 6. UMAP of Integrated Clusters
```{r}
DimPlot(obj, reduction = "umap.cca", group.by = "cca_clusters", label = TRUE)
library(dplyr)
library(tidyverse)
library(tibble)
cluster.sample <- table(obj$orig.ident, obj$cca_clusters) %>%
    as.data.frame.matrix() %>%
    rownames_to_column(var = "sample")
  cluster.sample[-1] <- lapply(cluster.sample[-1], function(x) x/sum(x))
  cluster.sample <- cluster.sample %>%
    pivot_longer(
      cols = -c("sample"),
      names_to = "cluster",
      values_to = "count"
    )
  cluster.sample$cluster <- factor(cluster.sample$cluster, levels = unique(cluster.sample$cluster))
ggplot(cluster.sample, aes(x=cluster, y=count, fill=sample)) +
    geom_bar(stat="identity") +
    theme_minimal()

DimPlot(
  obj,
  reduction = "umap.cca",
  group.by = "cca_clusters",
  label = TRUE
)
DimPlot(
  obj,
  reduction = "umap.cca",
  group.by = "predicted_doublets",
  label = TRUE
)
``` 

### 7. DotPlot with Key Markers
```{r}
neuron_markers <- c("Tubb3", "Elavl4", "Ret", "Phox2b", "Snap25", "Uchl1")
glia_markers <- c("Sox10", "Plp1", "Fabp7", "Erbb3", "Gas7")
DotPlot(obj, features = c(neuron_markers, glia_markers), scale.max = 50) + coord_flip()
``` 

### 8. Relative Cluster Proportions by Sample
```{r}
library(tidyverse)
cluster.sample <- table(obj$orig.ident, obj$cca_clusters) %>%
  as.data.frame.matrix() %>%
  rownames_to_column(var = "sample")
cluster.sample[-1] <- lapply(cluster.sample[-1], function(x) x / sum(x))
cluster.sample <- cluster.sample %>%
  pivot_longer(cols = -sample, names_to = "cluster", values_to = "count")

ggplot(cluster.sample, aes(x = cluster, y = count, fill = sample)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Relative Cluster Proportions", x = "Cluster", y = "Proportion")
```
